---
title: "Watershed delineation"
author: "Kelly Hondula"
date: "3/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("nhdR")
# library(nhdR)
# library(FedData)
# if (!require(devtools)) install.packages('devtools')
# devtools::install_github("giswqs/whiteboxR")
# library(whitebox)
# library(raster)
library(igraph)
library(sf)
library(fs)
library(dplyr)
```

# data 

Locations of watershed boundary dataset and data for this project

```{r}
data_dir_wbd <- "/nfs/khondula-data/wbd"
data_dir_coastalsheds <- "/nfs/khondula-data/coastalsheds"
```

Download and unzip watershed boundary dataset if not done already

```{r}
# zipfiles <- sprintf("https://prd-tnm.s3.amazonaws.com/StagedProducts/Hydrography/WBD/HU2/Shape/WBD_%s_HU2_Shape.zip", stringr::str_pad(1:22, width = 2, pad = "0"))
# destfiles <- sprintf(file.path(data_dir_wbd, "huc_%s.zip"), stringr::str_pad(1:22, width = 2, pad = "0"))
# purrr::walk2(.x = zipfiles, .y = destfiles, .f = ~download.file(url = .x, destfile = .y))
# unzipfiles <- sprintf(file.path(data_dir_wbd, "Shape_%s"), stringr::str_pad(1:22, width = 2, pad = "0"))
# purrr::walk2(.x = destfiles, .y = unzipfiles, .f = ~unzip(.x, exdir = .y))
```

HUC 2 geodatabase file downloaded from https://nrcs.app.box.com/v/gateway/folder/39640323180

```{r}
# download.file(url = "https://prd-tnm.s3.amazonaws.com/StagedProducts/Hydrography/WBD/National/GDB/WBD_National_GDB.zip", destfile = file.path(data_dir_wbd, "nationalwbd.zip"))
# unzip(file.path(data_dir_wbd, "nationalwbd.zip"), exdir = data_dir_wbd)
```

Read in ncaa points file created with consistent coordinate reference system and project to watershed boundary CRS. Note that SITE_ID column gets converted to SITE_ when writing as Esri shapefile

```{r}
ncaa_coords_sf <- st_read(file.path(data_dir_coastalsheds, "ncaa_coords", "ncaa_coords_wgs84.shp")) %>%
ncaa_coords_sf_prj <- st_transform(ncaa_coords_sf, 4269)
```

# Workflow step by step 

Select one site ID to test out workflow

```{r}
ncaa_siteID <- ncaa_coords_sf_prj$SITE_[4]
# ncaa_siteID <- "NCCAGL10-1026"
# ncaa_siteID <- "NCCAGL10-NPS09-096" # lake michigan example
my_ncaa_pt <- filter(ncaa_coords_sf_prj, SITE_ == ncaa_siteID)[1]
mypt_id <- ncaa_siteID
```

find which huc 2 the point is in, or the nearest HUC 2

```{r}
huc2 <- st_read(file.path(data_dir_wbd, "wbdhu2_a_us_september2018.gdb")) %>% st_transform(4269)
my_huc2_mat <- st_contains(huc2, my_ncaa_pt, sparse = FALSE)
my_huc2_id <- which(apply(my_huc2_mat, 1, any))
my_huc2 <- huc2[my_huc2_id,]
my_huc2_id <- my_huc2$HUC2 %>% as.character()
my_huc2_id
if(length(my_huc2_id)==0){
  my_huc2 <- huc2[which.min(st_distance(my_ncaa_pt, huc2)),]
  my_huc2_id <- my_huc2$HUC2 %>% as.character()
}
my_huc2_id
```

then read in HUC 12s for that watershed

```{r}
huc12_filepath <- file.path(data_dir_wbd, sprintf("Shape_%s/Shape/WBDHU12.shp", my_huc2_id))
huc12 <- st_read(huc12_filepath) %>% st_transform(4269)

my_huc12_mat <- st_contains(huc12, my_ncaa_pt, sparse = FALSE)
my_huc12_id <- which(apply(my_huc12_mat, 1, any))
my_huc12 <- huc12[my_huc12_id,]
my_huc12_id <- my_huc12$HUC12 %>% as.character()
my_huc12_id

if(length(my_huc12_id)==0){
  my_huc12 <- huc12[which.min(st_distance(my_ncaa_pt, huc12)),]
  my_huc12_id <- my_huc12$HUC12 %>% as.character()
}
```

Convert to an edgelist network to find all upstream HUC12s

```{r}
huc12_edgelist <- huc12 %>% st_drop_geometry() %>% dplyr::select(HUC12, ToHUC)
huc12_network <- huc12_edgelist %>% igraph::graph_from_data_frame()
paths_in <- igraph::all_simple_paths(huc12_network, from = my_huc12_id, mode = "in")
upstream_huc12s <- sapply(paths_in, names) %>% unlist() %>% unique()
upstream_huc12s
```

Filter those from huc 12 dataset, combine with huc 12 the point is in in case there are no upstream hucs, and union into one polygon

```{r}
upstream_huc12s_sf <- dplyr::filter(huc12, HUC12 %in% upstream_huc12s) %>% rbind(my_huc12)
upstream_huc12s_sf_union <- upstream_huc12s_sf %>% st_union()
```

Map
```{r}
leaflet() %>%
  addTiles() %>%
  # addPolygons(data = my_huc12) %>%
  addPolygons(data = upstream_huc12s_sf_union) %>%
  addMarkers(data = my_ncaa_pt)
```

save
```{r}
st_write(upstream_huc12s_sf, file.path(data_dir_coastalsheds, "ncaa_huc12sheds", 
                                         sprintf("%s_huc12shed.shp", mypt_id)))
st_write(upstream_huc12s_sf_union, file.path(data_dir_coastalsheds, "ncaa_huc12sheds_union",
                                               sprintf("%s_huc12shed_union.shp", mypt_id)))
```

# Workflow function

```{r}
ncaa_siteID <- ncaa_siteIDs_GL[1]
ncaa_siteID <- "NCCAGL10-GLBA10-150"
```

Self-contained function based on a site ID

```{r}
save_huc12shed_for_pt <- function(ncaa_siteID){
  # locations for data
  data_dir_wbd <- "/nfs/khondula-data/wbd"
  data_dir_coastalsheds <- "/nfs/khondula-data/coastalsheds"
  # where to save output shapefiles
  out_path <- file.path(data_dir_coastalsheds, "ncaa_huc12sheds")
  out_path2 <- file.path(data_dir_coastalsheds, "ncaa_huc12sheds_union")
  if(!fs::dir_exists(out_path)){fs::dir_create(out_path)}
  if(!fs::dir_exists(out_path2)){fs::dir_create(out_path2)}
  
  # read in ncaa coordinates
  ncaa_coords_sf_prj <- st_read(file.path(data_dir_coastalsheds, "ncaa_coords", "ncaa_coords_wgs84.shp")) %>%
    st_transform(4269) # project coords to WBD prj
  # select one point
  my_ncaa_pt <- dplyr::filter(ncaa_coords_sf_prj, SITE_ == ncaa_siteID) %>% slice(1)
  # mypt_id <- my_ncaa_pt$SITE_
  mypt_id <- ncaa_siteID
  
  # identify which huc 2 the point is in
  huc2 <- st_read(file.path(data_dir_wbd, "wbdhu2_a_us_september2018.gdb")) %>% st_transform(4269)
  my_huc2_mat <- st_contains(huc2, my_ncaa_pt, sparse = FALSE)
  my_huc2_id <- which(apply(my_huc2_mat, 1, any))
  my_huc2 <- huc2[my_huc2_id,]
  my_huc2_id <- my_huc2$HUC2 %>% as.character()
  # if point is not within any HUC2s then need to find closest
  if(length(my_huc2_id)==0){
    my_huc2 <- huc2[which.min(st_distance(my_ncaa_pt, huc2)),]
    my_huc2_id <- my_huc2$HUC2 %>% as.character()
  }
  
  # read in the corresponding huc 12s
  huc12_filepath <- file.path(data_dir_wbd, sprintf("Shape_%s/Shape/WBDHU12.shp", my_huc2_id))
  huc12 <- st_read(huc12_filepath) %>% st_transform(4269)
  
  # find which huc 12 point is in
  my_huc12_mat <- st_contains(huc12, my_ncaa_pt, sparse = FALSE)
  my_huc12_id <- which(apply(my_huc12_mat, 1, any))
  my_huc12 <- huc12[my_huc12_id,]
  my_huc12_id <- my_huc12$HUC12 %>% as.character()
  # if not within HUC12s then find closest
  if(length(my_huc12_id)==0){
    my_huc12 <- huc12[which.min(st_distance(my_ncaa_pt, huc12)),]
    my_huc12_id <- my_huc12$HUC12 %>% as.character()
  }
  
  # FOR GREAT LAKES
  # variable for whether point is in a great lake huc 12
  my_huc12_GL <- my_huc12_id %in% c("041800000300", # Lake superior
                                  "041800000100", # frontal lake superior
                                  "042400000300", # Lake Huron
                                  "040700030401", # frontal lake huron
                                  "041900000002", # Lake Michigan
                                  # "", # frontal lake michigan
                                  "042600000300", # Lake Erie
                                  "041502000200", # Lake Ontario
                                  "041502000100", # Frontal Lake Ontario
                                  "042600000200" # Frontal lake eries
                                  )
  if(my_huc12_id %in% c("042600000300", "042600000200")){
    # combine lake erie and frontal lake erie
    my_huc12 <- st_union(filter(huc12, HUC12 == "042600000300"), 
                       filter(huc12, HUC12 == "042600000200"))
  }
  if(my_huc12_id %in% c("041502000200", "041502000100")){
    # combine lake ontario and frontal lake ontario
     my_huc12 <- st_union(filter(huc12, HUC12 == "041502000200"), 
                       filter(huc12, HUC12 == "041502000100"))
  }
  if(my_huc12_id %in% c("041800000300", "041800000100")){
    # combine lake superior and frontal lake superior
    my_huc12 <- st_union(filter(huc12, HUC12 == "041800000300"), 
                       filter(huc12, HUC12 == "041800000100"))
  }

# if(my_huc12_id %in% c("041900000002", "")){
#   # combine lake michigan and frontal lake michigan
#   my_huc12 <- st_union(filter(huc12, HUC12 == "041900000002"), 
#                        filter(huc12, HUC12 == ""))
# }

  if(my_huc12_id %in% c("042400000300", "040700030401")){
    # combine lake huron and frontal lake huron
    my_huc12 <- st_union(filter(huc12, HUC12 == "042400000300"), 
                       filter(huc12, HUC12 == "040700030401"))
  }

  if(my_huc12_GL){
    # find the point on the outside of the 10m buffered lake polygon that is closest to the point
    my_huc12_ls <- my_huc12 %>% 
      st_transform(5070) %>% 
      st_buffer(500) %>% # try to get beyond the frontal watershed
      st_transform(4269) %>%
      st_cast("POLYGON") %>% st_cast("MULTILINESTRING") # convert to line
    shoreline_pt <- st_nearest_points(my_huc12_ls, my_ncaa_pt) %>% st_cast("POINT")
    shoreline_pt_sf <- st_sf(id = 1:2, shoreline_pt)[1,]
  
    my_huc12_mat <- st_contains(huc12, shoreline_pt_sf, sparse = FALSE)
    my_huc12_id <- which(apply(my_huc12_mat, 1, any))
    my_huc12 <- huc12[my_huc12_id,]
    my_huc12_id <- my_huc12$HUC12 %>% as.character()
      # if shoreline point is not in a huc 12, find closest one
    if(length(my_huc12_id)==0){
      my_huc12 <- huc12[which.min(st_distance(shoreline_pt_sf, huc12)),]
      my_huc12_id <- my_huc12$HUC12 %>% as.character()
    }
  }

  # convert to graph to find all upstream huc 12s
  huc12_edgelist <- huc12 %>% st_drop_geometry() %>% dplyr::select(HUC12, ToHUC)
  huc12_network <- huc12_edgelist %>% igraph::graph_from_data_frame()
  paths_in <- igraph::all_simple_paths(huc12_network, from = my_huc12_id, mode = "in")
  upstream_huc12s <- sapply(paths_in, names) %>% unlist() %>% unique()
  
  # filter from huc 12 object and save
  upstream_huc12s_sf <- dplyr::filter(huc12, HUC12 %in% upstream_huc12s) %>% 
    rbind(my_huc12) # combine with original huc12 in case no upstream ones found
  upstream_huc12s_sf_union <- upstream_huc12s_sf %>% st_union()
  
  st_write(upstream_huc12s_sf,
           file.path(data_dir_coastalsheds, "ncaa_huc12sheds", sprintf("%s_huc12shed.shp", mypt_id)),
            delete_dsn = TRUE) # overwrite if needed
  st_write(upstream_huc12s_sf_union, 
           file.path(data_dir_coastalsheds, "ncaa_huc12sheds_union", sprintf("%s_huc12shed_union.shp", mypt_id)),
           delete_dsn = TRUE) # overwrite if needed

}
```

Iterate over all points
```{r}
data_dir_coastalsheds <- "/nfs/khondula-data/coastalsheds"
ncaa_coords_sf_prj <- st_read(file.path(data_dir_coastalsheds, "ncaa_coords", "ncaa_coords_wgs84.shp")) %>%
    st_transform(4269) 
ncaa_siteIDs <- ncaa_coords_sf_prj$SITE_ %>% unique()
save_huc12shed_for_pt(ncaa_siteID = ncaa_siteIDs_GL[10])
# system.time(save_huc12shed_for_pt(ncaa_siteID = ncaa_siteIDs[3]))
# purrr::walk(ncaa_siteIDs, ~save_huc12shed_for_pt(.x))
```

```{r}
# find which points havent beendone yet
huc12shed_shpfiles <- fs::dir_ls(file.path(data_dir_coastalsheds, "ncaa_huc12sheds_union"), glob = "*.shp")
huc12shed_ids <- basename(huc12shed_shpfiles) %>% stringr::str_replace("_huc12shed_union.shp", "")
huc12shed_ids %>% length()
ncaa_siteIDs_toget <- ncaa_siteIDs[!ncaa_siteIDs %in% huc12shed_ids]
ncaa_siteIDs_GL <- grep("GL", ncaa_siteIDs, value = TRUE)
ncaa_siteIDs_toget %>% as.character()
```

Run on cluster
```{r}
library(rslurm)
pars <- data.frame(ncaa_siteID = ncaa_siteIDs_GL, stringsAsFactors = FALSE)

sjob2 <- rslurm::slurm_apply(save_huc12shed_for_pt, pars, 
                            jobname = "GL",
                            slurm_options = list(partition = "sesync", 
                                                 time = "12:00:00"),
                            nodes = 8, cpus_per_node = 2, 
                            submit = TRUE)

print_job_status(sjob2)
# rslurm::cancel_slurm(sjob2)
```

